An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-10 paper/paper.tex
Reformat as MQP paper

\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage[dvipdfm]{graphicx}
%\usepackage{bmpsize}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage[margin=1.15in]{geometry}

@@

\author{
  Bouffard, Daniel\\
  \texttt{bouffard.daniel@wpi.edu}
  \and
  Green, Marc\\
  \texttt{marcgreen@wpi.edu}
}

\title{An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems}

\date{\today}

\begin{document}
\maketitle

\section{Background Information}
\subsection{Centralized Backup Systems}
% TODO: get reliable sources for number of users of OneDrive and Google Drive
Several popular backup systems currently in use have a centralized system design. In these systems, there are many clients that all back up their data to a central location controlled by a single entity. Common elements of these systems include a fixed amount of data storage per person, with the option to obtain more for a fee. Two of the more popular systems are Dropbox, which has over 200 million users \cite{dropboxusers}, OneDrive, with over 250 million users, and Google Drive, with over 120 million users.

@@

\subsubsection{Overlay Networks}

@@

\subsection{Peer to Peer Backup Systems}
Peer to Peer Backup Systems are networks that allow users to backup data without the use of a central entity. Instead, each user offers some of his or her storage space in order to use the space of others as backup locations. There are several common components of existing P2P backup systems which are described in the following sections.

\subsubsection{File Chunking}
The format in which files are stored can have an effect on the efficiency of the system. pStore and PeerStore split files up into chunks, rather than distributing files to peers in their entirety. They also use special files, called File Block Lists in pStore and Descriptors in PeerStore, that list which blocks belong to the file and where to find them in the network. This has the benefit of having data units be a fixed size that will be easy to store for any node. For example, if a peer needs to store a 10GB file, it does not need to find a peer capable of storing 10GB (which might be rare); rather, it needs to find $file\_size/chunk\_size*number\_of\_replicas$ peers. Although lowering the chunk size will increase the chances of finding a peer to store on, it is important to choose a reasonable chunk size, as using more nodes for storage will add more overhead in finding those nodes.

@@

\subsubsection{Fairness}
Another aspect of P2P backup systems is fairness between peers. Peers need to be storing an amount of data proportional to the amount that they are backing up to other peers. This can be challenging, because untrustworthy peers can attempt to game the system to obtain free backup storage. For example, consider a simple system where each peer must offer store the same amount of data that they back up on other peers. A peer could claim that all of the storage space that it should be offering is being used by other peers, while in reality the peer has not taken in any peer's data. It is clear that a mechanism needs to be built into the system that ensures that peers store the data of each other. PeerStore handles this via a combined trading and challenge system. When a peer wants to create a backup of its data, it finds a peer to trade data with. In its request to backup data, the peer will also advertise that it is willing to store a certain amount of data. The peer that receives the request will then need to accept or reject it, and if the request is accepted the receiver will also need to request a certain amount of storage. If the original requestor accepts the offer, then the two become trading partners and will send each other data \cite{PeerStore}.

@@

\subsubsection{Churn and Data Migration}
With each storage node being regular users, it is often the case that nodes can be offline. \textit{Node churn} refers to fact that peers will continuously enter and leave the P2P network \cite{StorageSearchP2PNetworks}. The designer of a P2P backup system must take this into consideration if the system is to provide reliable access to backed up data. This has been accomplished via data migration \cite{pStore,PeerStore}, where the availability of files is periodically checked and replicas are made when the nodes being used for backup are not reliably available. Other data migration algorithms have been designed specifically to work in high churn networks \cite{StorageSearchP2PNetworks}. In this case, data replicas are periodically passed to different groups of nodes, which means that newer nodes have a chance to store data. This is beneficial because data gets evenly distributed throughout the overlay network, so any given node is not more important to the network than any others with regard to the amount of data it is storing.

@@

\subsubsection{Cryptography}
% TODO: the figure for convergent encryption might be nice here
Since data is being stored on several untrusted peers, each user's data need to be encrypted before backup. A technique used in pStore and PeerStore is convergent encryption \cite{pStore, PeerStore}. To use this encryption mechanism, the hash of the data to encrypt is calculated, and then the hash is used as the symmetric key to encrypt the data. The main benefit of using convergent encryption is that the same data will be encrypted the same way each time, so identical blocks in the network can be shared between peers in their encrypted form.

@@

While this is a convenient optimization, it allows for users who have access to the unencrypted version of a file to confirm whether or not that file is being stored, since no more information is needed to encrypt or decrypt a given file.
%\subsubsection{Case Study: pStore}
%\subsubsection{Case Study: PeerStore}
\subsection{Distributed Hash Tables}

@@

\subsection{The BitTorrent Protocol}

@@

\subsection{BitTorrent Sync}
BitTorrent Sync is a new application developed by BitTorrent, Inc. It is a P2P synchronization tool that allows users to share files between trusted devices. This is accomplished by generating a set of 20-byte "secrets" for each folder to share and distributing the secret to each trusted device. All data is encrypted using AES-128 in counter mode \cite{btsynctech}, the key of which is derived from the secret. There are three different secrets, each of which provide different permissions for the peer that is given the secret: the read/write secret, the read only secret, and the encryption secret, which allows the node to store an encrypted version of the data, so that it cannot read the contents. A variation of the read/write and read only secrets also exists called the one-time secret, which allows a peer to send out a valid key that is only usable by one node \cite{btsyncuserguide}.

@@

\section{Implementation}
\subsection{Languages and Tools}
\subsection{System Design}
\subsubsection{Overview}
\subsubsection{}
\subsubsection{DHT}

>>>>

\documentclass[12pt]{report}

%%\usepackage{setspace}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{alltt}
\usepackage{color}
\usepackage{fancyvrb} 
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{url}
\usepackage[autostyle]{csquotes} 
\usepackage[toc,page]{appendix}

%\usepackage[top=1in, bottom=1.5in, left=1.5in,right=1.5in]{geometry} % setting the page alignment with this package

@@

\begin{document}
\newpage
\begin{center}
\thispagestyle{empty}
\Large{\textbf{A Major Qualifying Project Report\\ \large{ON}}}\\[0.7cm]
\LARGE{\textsc {\textbf{An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems}}}\\[0.5cm]
\vspace{0.2cm}
\Large{\textbf{\\Submitted to the Faculty of }}
\LARGE{\textbf{\\WORCESTER POLYTECHNIC INSTITUTE\\}}
\vspace{0.5cm}
\Large{\textbf{\\In Partial Fulfillment of the Requirement for the}}
\Large{\textbf{\\Degree of Bachelor of Science}}
\vspace{0.5cm}
\Large{\textbf{\\by}}\\[0.5cm]
\large{\textbf{Dan Bouffard}}\\
\large{\textbf{Marc Green}}\\
\vspace{0.5cm}
\large{\textbf{UNDER THE GUIDANCE OF}}\\
\large{\textbf{Professor G\'abor N. S\'ark\"ozy}}\\
\vspace{1cm}
\large{\textbf{\\May 1, 2014}}\\
\newpage
\end{center}

\begin{abstract}
Abstract
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
Acknowledgements
\end{abstract}

\tableofcontents
\listoffigures

\chapter{Introduction}

Introduction

\chapter{Background Information}
\section{Centralized Backup Systems}
% TODO: get reliable sources for number of users of OneDrive and Google Drive
Several popular backup systems currently in use have a centralized system design. In these systems, there are many clients that all back up their data to a central location controlled by a single entity. Common elements of these systems include a fixed amount of data storage per person, with the option to obtain more for a fee. Two of the more popular systems are Dropbox, which has over 200 million users \cite{dropboxusers}, OneDrive, with over 250 million users, and Google Drive, with over 120 million users.

@@

\subsection{Overlay Networks}

@@

\section{Peer to Peer Backup Systems}
Peer to Peer Backup Systems are networks that allow users to backup data without the use of a central entity. Instead, each user offers some of his or her storage space in order to use the space of others as backup locations. There are several common components of existing P2P backup systems which are described in the following sections.

\subsection{File Chunking}
The format in which files are stored can have an effect on the efficiency of the system. pStore and PeerStore split files up into chunks, rather than distributing files to peers in their entirety. They also use special files, called File Block Lists in pStore and Descriptors in PeerStore, that list which blocks belong to the file and where to find them in the network. This has the benefit of having data units be a fixed size that will be easy to store for any node. For example, if a peer needs to store a 10GB file, it does not need to find a peer capable of storing 10GB (which might be rare); rather, it needs to find $file\_size/chunk\_size*number\_of\_replicas$ peers. Although lowering the chunk size will increase the chances of finding a peer to store on, it is important to choose a reasonable chunk size, as using more nodes for storage will add more overhead in finding those nodes.

@@

\subsection{Fairness}
Another aspect of P2P backup systems is fairness between peers. Peers need to be storing an amount of data proportional to the amount that they are backing up to other peers. This can be challenging, because untrustworthy peers can attempt to game the system to obtain free backup storage. For example, consider a simple system where each peer must offer store the same amount of data that they back up on other peers. A peer could claim that all of the storage space that it should be offering is being used by other peers, while in reality the peer has not taken in any peer's data. It is clear that a mechanism needs to be built into the system that ensures that peers store the data of each other. PeerStore handles this via a combined trading and challenge system. When a peer wants to create a backup of its data, it finds a peer to trade data with. In its request to backup data, the peer will also advertise that it is willing to store a certain amount of data. The peer that receives the request will then need to accept or reject it, and if the request is accepted the receiver will also need to request a certain amount of storage. If the original requestor accepts the offer, then the two become trading partners and will send each other data \cite{PeerStore}.

@@

\subsection{Churn and Data Migration}
With each storage node being regular users, it is often the case that nodes can be offline. \textit{Node churn} refers to fact that peers will continuously enter and leave the P2P network \cite{StorageSearchP2PNetworks}. The designer of a P2P backup system must take this into consideration if the system is to provide reliable access to backed up data. This has been accomplished via data migration \cite{pStore,PeerStore}, where the availability of files is periodically checked and replicas are made when the nodes being used for backup are not reliably available. Other data migration algorithms have been designed specifically to work in high churn networks \cite{StorageSearchP2PNetworks}. In this case, data replicas are periodically passed to different groups of nodes, which means that newer nodes have a chance to store data. This is beneficial because data gets evenly distributed throughout the overlay network, so any given node is not more important to the network than any others with regard to the amount of data it is storing.

@@

\subsection{Cryptography}
% TODO: the figure for convergent encryption might be nice here
Since data is being stored on several untrusted peers, each user's data need to be encrypted before backup. A technique used in pStore and PeerStore is convergent encryption \cite{pStore, PeerStore}. To use this encryption mechanism, the hash of the data to encrypt is calculated, and then the hash is used as the symmetric key to encrypt the data. The main benefit of using convergent encryption is that the same data will be encrypted the same way each time, so identical blocks in the network can be shared between peers in their encrypted form.

@@

While this is a convenient optimization, it allows for users who have access to the unencrypted version of a file to confirm whether or not that file is being stored, since no more information is needed to encrypt or decrypt a given file.
%\subsection{Case Study: pStore}
%\subsection{Case Study: PeerStore}
\section{Distributed Hash Tables}

@@

\section{The BitTorrent Protocol}

@@

\section{BitTorrent Sync}
BitTorrent Sync is a new application developed by BitTorrent, Inc. It is a P2P synchronization tool that allows users to share files between trusted devices. This is accomplished by generating a set of 20-byte "secrets" for each folder to share and distributing the secret to each trusted device. All data is encrypted using AES-128 in counter mode \cite{btsynctech}, the key of which is derived from the secret. There are three different secrets, each of which provide different permissions for the peer that is given the secret: the read/write secret, the read only secret, and the encryption secret, which allows the node to store an encrypted version of the data, so that it cannot read the contents. A variation of the read/write and read only secrets also exists called the one-time secret, which allows a peer to send out a valid key that is only usable by one node \cite{btsyncuserguide}.

@@

\chapter{Implementation}
\section{Languages and Tools}
\section{System Design}
\subsection{Overview}
\subsection{}
\subsection{DHT}
